---
title: "Linear regression in SCP"
subtitle: "Resolving the issue of msqrob2"
author: "Léopold Guyot"
institute: "UGent"
date: "24/07/2024"
output:
  xaringan::moon_reader:
    css: styles.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false

---

class:

```{r xaringanExtra, echo=FALSE}
xaringanExtra::use_tile_view()  ## use o to get an overview
xaringanExtra::use_fit_screen() ## use Alt-F to fit to screen
xaringanExtra::use_extra_styles(
  hover_code_line = TRUE
  )
xaringanExtra::use_panelset()
```

``` {r packages, include=FALSE}

library(QFeatures)
library(msqrob2)
library(limma)
library(scp)
library(BiocParallel)
library(tidyverse)

```


# Data loading

.panelset[
.panel[
.panel-name[Code]

```{r data_loading}

.createMockBalanced <- function(k, runs) {
    out <- split(runs, runs)
    out <- lapply(out, function(x) {
        index <- 1:length(x)
        x[sample(index)] <- rep(1:k, length.out = length(index))
        x
    })
    out <- S4Vectors::unname(do.call(c, out))
    as.factor(out)
}

sce <- readRDS("data/leduc_mock.rds")
set.seed(1234)
sce$Mock <- .createMockBalanced(2, sce$Set)

```
]

.panel[
.panel-name[Preview]

```{r sce_preview}
sce
```
]
]


---
class:middle

# Subset of specific peptides

### Non fitted

```{r nfit}
non_fitted <- sce["AAATAEEPDPK_2",]
```
### Fitted with each method

```{r fit}
fitted <- sce["GAEAANVTGPGGVPVQGSK_3",] # find a new, msqrob not fitting
```

### Non fitted with msqrob2

```{r nfit_msqrob}
non_fitted_msqrob <- sce["EATTEFSVDAR_2",]
```


---
class: 

# Modified version of msqrobLm
.panelset[
.panel[
.panel-name[Comments]
- The issue with msqrobLm is that it does not handle the case where the model
is rank deficient.
- msqrobLm already include a subsetting of the design matrix to remove columns
that have no information (i.e. all zeros). But it does not handle the case
where the reference group of a variable has no values, in this case the model
become rank deficient.
- To solve this issue, we changed the way the subset is performed by only
keeping the x first columns of the design matrix, with columns sorted by their
pivoting rank and x being the rank of the qr decomposition of the design matrix.
]
.panel[
.panel-name[Code]

.scrollable[
```{r msqrobLm_modif}

msqrob2_fix <- function (y, formula, data, robust = TRUE, maxitRob = 5) 
{
    myDesign <- model.matrix(formula, data)
    models <- apply(y, 1, function(y, design) {
        obs <- is.finite(y)
        type <- "fitError"
        model <- list(coefficients = NA, vcovUnscaled = NA, 
            sigma = NA, df.residual = NA, w = NA)
        if (sum(obs) > 0) {
            X <- design[obs, , drop = FALSE]
            qrX <- qr(X)
            X <- X[, qrX$pivot[seq_len(qrX$rank)], drop = FALSE]
            # => old version
            # X <- X[, colMeans(X == 0) != 1, drop = FALSE] 
            y <- y[obs]
            colnames_orig <- colnames(design)
            if (robust) {
                mod <- try(MASS::rlm(X, y, method = "M", maxit = maxitRob), 
                    silent = TRUE)
                if (!is(mod, "try-error")) {
                    type <- "rlm"
                }
            }
            else {
                mod <- try(lm.fit(X, y))
                if ((!is(mod, "try-error")) & mod$rank == ncol(X)) {
                    type <- "lm"
                }
            }
            if (type == "rlm") {
                w <- mod$w
                sigma <- sqrt(sum(mod$w * mod$resid^2)/(sum(mod$w) - 
                    mod$rank))
                df.residual <- sum(mod$w) - mod$rank
                if (df.residual < 2L) 
                    type <- "fitError"
            }
            if (type == "lm") {
                w <- NULL
                sigma <- sqrt(sum(mod$residuals^2/mod$df.residual))
                df.residual <- mod$df.residual
                if (df.residual < 2L) 
                    type <- "fitError"
            }
            if (type != "fitError") {
                coef <- rep(NA, length(colnames_orig))
                names(coef) <- colnames_orig
                coef[names(mod$coef)] <- mod$coef
                vcovUnscaled <- matrix(NA, nrow = length(colnames_orig), 
                    ncol = length(colnames_orig))
                rownames(vcovUnscaled) <- colnames(vcovUnscaled) <- colnames_orig
                vcovUnscaled[names(mod$coef), names(mod$coef)] <- msqrob2:::.vcovUnscaled(mod)
                model <- list(coefficients = mod$coef, vcovUnscaled = msqrob2:::.vcovUnscaled(mod), 
                    sigma = sigma, df.residual = df.residual, 
                    w = w)
            }
        }
        .StatModel(type = type, params = model, varPosterior = as.numeric(NA), 
            dfPosterior = as.numeric(NA))
    }, design = myDesign)
    hlp <- limma::squeezeVar(var = vapply(models, getVar, numeric(1)), 
        df = vapply(models, getDF, numeric(1)))
    for (i in seq_len(length(models))) {
        mydf <- hlp$df.prior + getDF(models[[i]])
        models[[i]]@varPosterior <- as.numeric(hlp$var.post[i])
        models[[i]]@dfPosterior <- as.numeric(mydf)
    }
    return(models)
}

```

]
]
]

---
class:

# Modelling

.panelset[
.panel[
.panel-name[Non fitted]

```{r modelling_nfit, results = 'hide', warning=FALSE}
scp_nfit <- scp::scpModelWorkflow(non_fitted,
                             formula = ~ 1 + Channel + Set + Mock)
msqrobLm_nfit <- msqrob2::msqrobLm(assay(non_fitted),
                                   ~ 1 + Channel + Set + Mock,
                                   colData(non_fitted),
                                   robust = FALSE)
msqrobLm_fix_nfit <- msqrob2_fix(assay(non_fitted),
                           ~ 1 + Channel + Set + Mock,
                           colData(non_fitted),
                           robust = FALSE)
y <- assay(non_fitted)
design <- model.matrix(~ 1 + Channel + Set + Mock, colData(non_fitted))
limma_nfit <- limma::lmFit(y, design)
```
]

.panel[
.panel-name[Fitted with each method]


```{r modelling_fit, results = 'hide', warning=FALSE}
scp_fit <- scp::scpModelWorkflow(fitted,
                            formula = ~ 1 + Channel + Set + Mock)
msqrobLm_fit <- msqrob2::msqrobLm(assay(fitted),
                                  ~ 1 + Channel + Set + Mock,
                                  colData(fitted),
                                  robust = FALSE)
msqrobLm_fix_fit <- msqrob2_fix(assay(fitted),
                            ~ 1 + Channel + Set + Mock,
                            colData(fitted),
                            robust = FALSE)
y <- assay(fitted)
design <- model.matrix(~ 1 + Channel + Set + Mock, colData(fitted))
limma_fit <- limma::lmFit(y, design)
```

]
.panel[
.panel-name[Non fitted with msqrobLm]


```{r modelling_fit_msqrob, results = 'hide', warning=FALSE}
scp_nfit_msqrob <- scp::scpModelWorkflow(non_fitted_msqrob,
                                          formula = ~ 1 + Channel + Set + Mock)
msqrobLm_nfit_msqrob <- msqrob2::msqrobLm(assay(non_fitted_msqrob),
                                          ~ 1 + Channel + Set + Mock,
                                          colData(non_fitted_msqrob),
                                          robust = FALSE)
msqrobLm_fix_nfit_msqrob <- msqrob2_fix(assay(non_fitted_msqrob),
                                        ~ 1 + Channel + Set + Mock,
                                        colData(non_fitted_msqrob),
                                        robust = FALSE)
y <- assay(non_fitted_msqrob)
design <- model.matrix(~ 1 + Channel + Set + Mock, colData(non_fitted_msqrob))
limma_nfit_msqrob <- limma::lmFit(y, design)
```

]


]
---

# Results non fitted

.panelset[
.panel[
.panel-name[Success results]
.scrollable[
```{r results_nfit}

result_scp_nfit <- scpModelFilterNPRatio(scp_nfit, filtered = FALSE)
result_msqrobLm_nfit <- msqrobLm_nfit[[1]]@type
result_msqrobLm_fix_nfit <- msqrobLm_fix_nfit[[1]]@type
result_limma_nfit <- sum(!is.na(limma_nfit$coefficients))

print(paste("The scplainer model return a NP ratio that is equal to",
            result_scp_nfit))
print("Some coefficients are still estimated due to theuse of a ridge regression.")
print(paste("The msqrobLm model return a",
            result_msqrobLm_nfit,
            "model."))
print(paste("The msqrobLm model with the fix return a"
            , result_msqrobLm_fix_nfit,
            "model."))
print(paste("The limma model return",
            result_limma_nfit,
            "coefficients."))
```
]]

.panel[
.panel-name[Comments]
- The scplainer model return a NP ratio that is less than 1, which means that 
this model is not usable in the scplainer workflow. But some coefficients are
still estimated due to the use of a ridge regression.
- The msqrobLm model return a fitError model, which means that the model is not
estimable
- The msqrobLm model with the fix return a fitError model, which means that 
the model is not estimable
- The limma model return some coefficients, which means that the model still
return a normal output even with poor data
]]

---

# Results fitted with each method

.panelset[
.panel[
.panel-name[Success results]
.scrollable[
```{r results_fit}

result_scp_fit <- scp::scpModelFilterNPRatio(scp_fit, filtered = FALSE)
result_msqrobLm_fit <- msqrobLm_fit[[1]]@type
result_msqrobLm_fix_fit <- msqrobLm_fix_fit[[1]]@type
result_limma_fit <- sum(!is.na(limma_fit$coefficients))

print(paste("The scplainer model return a NP ratio that is equal to",
            result_scp_fit))
print(paste("The msqrobLm model return a",
            result_msqrobLm_fit,
            "model."))
print(paste("The msqrobLm model with the fix return a"
            , result_msqrobLm_fix_fit,
            "model."))
print(paste("The limma model return",
            result_limma_fit,
            "coefficients."))
```
]]
.panel[
.panel-name[Coefficients results]
.scrollable[

```{r coef_fit}
coef_scp_fit <- scp:::scpModelCoefficients(scp_fit)[[1]]
coef_msqrobLm_fit <- getCoef(msqrobLm_fit[[1]])
coef_msqrobLm_fix_fit <- getCoef(msqrobLm_fix_fit[[1]])
coef_limma_fit <- limma_fit$coefficients
```
]]

.panel[
.panel-name[Comments]

]]

---

# Results non fitted with msqrob2

.panelset[
.panel[
.panel-name[Success results]
.scrollable[
```{r results_nfit_msqrob}

result_scp_nfit_msqrob <- scp::scpModelFilterNPRatio(scp_nfit_msqrob, filtered = FALSE)
result_msqrobLm_nfit_msqrob <- msqrobLm_nfit_msqrob[[1]]@type
result_msqrobLm_fix_nfit_msqrob <- msqrobLm_fix_nfit_msqrob[[1]]@type
result_limma_nfit_msqrob <- sum(!is.na(limma_nfit_msqrob$coefficients))

print(paste("The scplainer model return a NP ratio that is equal to",
            result_scp_nfit_msqrob))
print(paste("The msqrobLm model return a",
            result_msqrobLm_nfit_msqrob,
            "model."))
print(paste("The msqrobLm model with the fix return a"
            , result_msqrobLm_fix_nfit_msqrob,
            "model."))
print(paste("The limma model return",
            result_limma_nfit_msqrob,
            "coefficients."))
```
]]
.panel[
.panel-name[Coefficients results]
.scrollable[

```{r coef_nfit_msqrob}
coef_scp_nfit_msqrob <- scp:::scpModelCoefficients(scp_nfit_msqrob)[[1]]
coef_msqrobLm_nfit_msqrob <- getCoef(msqrobLm_nfit_msqrob[[1]])
coef_msqrobLm_fix_nfit_msqrob <- getCoef(msqrobLm_fix_nfit_msqrob[[1]])
coef_limma_nfit_msqrob <- limma_nfit_msqrob$coefficients

summary_coef_nfit_msqrob <- full_join(data.frame("parameter" = names(coef_msqrobLm_fix_nfit_msqrob),
                    "coefficient_msqrobLm_fix" = coef_msqrobLm_fix_nfit_msqrob),
      data.frame("parameter" = colnames(coef_limma_nfit_msqrob),
                    "coefficient_limma" = as.numeric(coef_limma_nfit_msqrob)), by = "parameter")
plot(x = summary_coef_nfit_msqrob$coefficient_msqrobLm_fix, y = summary_coef_nfit_msqrob$coefficient_limma)
summary_coef_nfit_msqrob <- pivot_longer(summary_coef_nfit_msqrob,
                                         cols = c("coefficient_msqrobLm_fix", "coefficient_limma"),
                                         names_to = "model",
                                         values_to = "coefficient")


```
]]

.panel[
.panel-name[Comments]

]]
